Human Factors Influencing Visual Statistical Inference

### 2015-08-28 sociology_paper.Rnw
up to Location Effect

Table~\ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt. Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated. Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments. 
The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments. The negative estimates suggest that on average later attempts took less time. Even though observers did not improve in their performance in successive attempts, they became more efficient in their response. The parameter $\alpha_1$ for first attempt is also highly significant. The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts. Most observers, before evaluating the first lineup, carefully read through instructions and familiarize themselves  with the types of plots used in the experiment, which would inflate the time taken on the first lineup. Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers. For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).  Later pages of the web site were identical  except for the lineup. Thus in the later attempts an observer does not need to spend any time for reading instructions. The model reflects that fact.

\begin{table}[htbp]
\centering
\caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup. Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7. }
\scalebox{0.81}{
\begin{tabular}{rrrrccrrrccrrrc}
  \hline
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\

@@

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
 \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model \eqref{eqn:trend_time} without covariate attempt. Differences in subject specific slopes are observed. Some of the subjects did worse over successive attempts while others did better. Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends. For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.}
   \label{fig:learning_trend_time}
\end{figure}

>>>>

Table~\ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt. Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated. Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments. 
The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments. The negative estimates suggest that on average later attempts took less time. Even though observers did not improve in their performance in successive attempts, they became more efficient in their response. The parameter $\alpha_1$ for first attempt is also highly significant. The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts. One explanation might be that participants take the chance to  carefully read through instructions and familiarize themselves  with the types of plots used in the experiment before evaluating their first lineup, which would inflate the time taken on the first lineup. Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers. For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).  Later pages of the web site were identical  except for the lineup. %Thus in the later attempts an observer does not need to spend any time for reading instructions. The model reflects that fact.

\begin{table}[htbp]
\centering
\caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup. Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7. }
\scalebox{0.81}{
\begin{tabular}{rrrrccrrrccrrrc}
  \hline
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\

@@

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
 \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model \eqref{eqn:trend_time} without covariate attempt. Differences in subject specific slopes are observed. Some of the subjects did worse over successive attempts while others did better. Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends. For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
   \label{fig:learning_trend_time}
\end{figure}
